Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-18 doc/manuscript/translocation.qmd
Minor edits in response to coauthor RG comments

Roland A. Knapp^a,b,^\*, Mark Q. Wilber^c,1^, Allison Q. Byrne^d,e,1^, Maxwell B. Joseph^f,g^, Thomas C. Smith^a,b^, Andrew P. Rothstein^d,e^, Rob L. Grasso^h^, Erica Bree Rosenblum^d,e^

@@

^h^Resources Management and Science, Yosemite National Park, El Portal, CA, 95318

@@

<!--#  Preprint Server: bioRxiv  -->

@@

Despite the resistance/tolerance of adult MYL frogs in recovering populations, individual and population-level impacts of Bd are still evident.
In an earlier study of 2 of our 12 translocated populations [@joseph2018], Bd infection and load had detectable effects on the survival of adults and may have influenced population establishment (sites referred to as "Alpine" and "Subalpine" in [@joseph2018] are identified as "70550" and "70505" in the current study).
Applying similar analyses to all 12 of our translocated populations would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on adults, MYL frogs immediately following metamorphosis ("metamorphs") are highly susceptible to Bd infection [@ellison2018] and experience high mortality.
This high susceptibility of metamorphs is documented for numerous species of anurans, and may result from the poorly developed immune system characteristic of this early life stage [@humphries2022].
In naturally recovering and translocated MYL frog populations, we suggest that the high mortality of metamorphs is an important limitation on subsequent recruitment of new adults.
Therefore, although adult MYL frogs appear relatively resistant/tolerant, Bd infection continues to have important limiting effects on recovering populations [see also @hollanders2022].

@@

During each site visit, we conducted VES either immediately before CMR surveys or during the first day of CMR surveys.
VES was conducted by walking the entire water body perimeter, first 100 m of each inlet and outlet stream, and any fringing ponds and wetlands, and counting all *R. sierrae* tadpoles and juveniles.
These *R. sierrae* life stages have high detectablility, and counts are highly repeatable and provide estimates of relative abundance [@knapp2000].

>>>>

Roland A. Knapp^a,b,^\*, Mark Q. Wilber^c,1^, Allison Q. Byrne^d,e,1^, Maxwell B. Joseph^f,g^, Thomas C. Smith^a,b^, Andrew P. Rothstein^d,e,h^, Robert L. Grasso^i^, Erica Bree Rosenblum^d,e^

@@

^h^Current affiliation: Ginkgo Bioworks, Boston, MA, 02210

^i^Resources Management and Science, Yosemite National Park, El Portal, CA, 95318

@@

@@

Despite the resistance/tolerance of adult MYL frogs in recovering populations, individual and population-level impacts of Bd are still evident.
In an earlier study of 2 of our 12 translocated populations [@joseph2018], Bd infection and load had detectable effects on the survival of adults and may have influenced population establishment (sites referred to as "Alpine" and "Subalpine" in [@joseph2018] are identified as "70550" and "70505" in the current study).
Applying similar analyses to all 12 of our translocated populations would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on adults, the impacts on younger life stages are more apparent.
MYL frogs immediately following metamorphosis ("metamorphs") are highly susceptible to Bd infection [@ellison2018] and as a result experience high mortality.
This high susceptibility of metamorphs is documented for numerous species of anurans, and may result from the poorly developed immune system characteristic of this life stage [@humphries2022].
In naturally recovering and translocated MYL frog populations, we suggest that the high mortality of metamorphs is an important limitation on subsequent recruitment of new adults.
Therefore, although adult MYL frogs appear relatively resistant/tolerant, Bd infection continues to have important limiting effects on recovering populations [see also @hollanders2022].

@@

During each site visit, we conducted VES either immediately before CMR surveys or during the first day of CMR surveys.
VES was conducted by walking the entire water body perimeter, first 100 m of each inlet and outlet stream, and any fringing ponds and wetlands, and counting all *R. sierrae* tadpoles and juveniles.
These *R. sierrae* life stages have high detectability, and counts are highly repeatable and provide estimates of relative abundance [@knapp2000].
